---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# BiomassWorkflow

The goal of BiomassWorkflow is to streamline data analysis and removing the potential for errors in the process. The package is most useful for generating biomass data and comparing biomass between various groups. There are also a couple of helper functions included and other scripts that help with moving samples through the processing pipeline.

## Installation

You can install BiomassWorkflow from github with:

```{r gh-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("econtijoch/Biomass-Workflow")
```

If you need to install devtools first, you'll need to do the following first to download and update devtools to the most current version:

```{r devtools-installation, eval = FALSE}
install.packages("devtools")
devtools::install_github("hadley/devtools")
```

## Tutorial

I am working on updating the documentation for the tutorial. I'll get around to it soon.
